Transmembrane helices are also an overlooked source of major histocompatibility complex class II epitopes and evolutionary more conserved than expected by chance

### 2020-12-16 bbbq_article.tex
Quote dpSNP correctly

To do so, we use the 
NCBI (\url{https://www.ncbi.nlm.nih.gov/}) databases:
'gene' to find the gene names of membrane proteins, 
'dbSNP' (\cite{sachidanandam2001map}) for SNPs associated with those genes
and 'protein' to obtain the sequence of proteins that SNPs act upon.
The 'dbSNP' contains 2 billion (human-only) SNPs, although
it is estimated that approximately ten percent (ranging from 2 to 36 percent) 
are artifacts (\cite{carlson2003additional, cutler2001high, gabriel2002structure, mitchell2004discrepancies, musumeci2010single, reich2003quality}).

@@

The first query is a call to the 'gene' database for the 
term 'membrane protein' (in all fields) for the organism Homo sapiens.
This results in \richel{1126} gene IDs.
The next query is a call to the 'gene' database 
to obtain the gene names from the gene IDs.
Per gene name, the 'dbSNP' NCBI database is queried for 
SNPs associated with the gene name. The number of SNPs
was limited to the first \richel{100} SNPs per gene,
resulting in \richel{112600} SNPs.
Per selected SNP, all variation (i.e. descriptions of the changes on DNA
and/or RNA and/or protein level) are obtained from the 'dbSNP' database
as well. Of these variations, only the variations that result in a single
amino acid substitution are kept, resulting in \richel{???} protein changes.
Per protein substitution (i.e. a protein name and an index of the changed residue), 
the 'protein' NCBI database was queried for the
protein sequence.
Of each protein sequence, the protein topology is determined 
using PureseqTM.
From the topology and the known location of the amino acid substitution, 
it is tallied if the change occurred within a TMH or not.
Also, the percentage of the topology that is TMH is measured.
Not all membrane proteins have a transmembrane helix, resulting
in \richel{???} proteins with at least one TMH.

>>>>

To do so, we use the 
NCBI (\url{https://www.ncbi.nlm.nih.gov/}) databases:
'gene' to find the gene names of membrane proteins, 
'dbSNP' (\cite{sherry2001dbsnp}) for SNPs associated with those genes
and 'protein' to obtain the sequence of proteins that SNPs act upon.
The 'dbSNP' contains 2 billion (human-only) SNPs, although
it is estimated that approximately ten percent (ranging from 2 to 36 percent) 
are artifacts (\cite{carlson2003additional, cutler2001high, gabriel2002structure, mitchell2004discrepancies, musumeci2010single, reich2003quality}).

@@

The first query is a call to the 'gene' database for the 
term 'membrane protein' (in all fields) for the organism Homo sapiens.
This results in \richel{1126} gene IDs.
The next query is a call to the 'gene' database 
to obtain the gene names from the gene IDs.
Per gene name, the 'dbSNP' NCBI database is queried for 
SNPs associated with the gene name. The number of SNPs
was limited to the first \richel{100} SNPs per gene,
resulting in \richel{112600} SNPs 
\richel{Note that \cite{musumeci2010single} demonstrates for a way to get
try SNPs, instead of SNDs, maybe I should use that}
.
Per selected SNP, all variation (i.e. descriptions of the changes on DNA
and/or RNA and/or protein level) are obtained from the 'dbSNP' database
as well. Of these variations, only the variations that result in a single
amino acid substitution are kept, resulting in \richel{???} protein changes.
Per protein substitution (i.e. a protein name and an index of the changed residue), 
the 'protein' NCBI database was queried for the
protein sequence.
Of each protein sequence, the protein topology is determined 
using PureseqTM.
From the topology and the known location of the amino acid substitution, 
it is tallied if the change occurred within a TMH or not.
Also, the percentage of the topology that is TMH is measured.
Not all membrane proteins have a transmembrane helix, resulting
in \richel{???} proteins with at least one TMH.
